---
title: "testGEODatasets"
author: "Alex Sanchez"
html:
    toc: true
    theme: cosmo
    html-math-method: katex
    css: styles.css
---

```{r setup, echo=FALSE}
library(knitr)
library(rmdformats)

## Global options
options(max.print="75")
opts_chunk$set(echo=TRUE,
	             cache=FALSE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```

## Presentation

File `GEOdatasets.xls` contains a of references to Gene Expression Omnibus datasets that can be used for performing microarray analysis.

```{r}
library(readxl)
GEOdatasets <- read_excel("GEOdatasets_updatedNov2022.xls")
colnames(GEOdatasets)
Series <- GEOdatasets["GEO Serie"]
Datasets <- GEOdatasets["GEO Dataset"]
```

While thay have been tested along time it is good to check their availability and characteristics, which is what th current document does.

```{r}
library(Biobase)
library(GEOquery)
if(!file.exists("GEOSeries.Rda")){
  gseList <- list()
  for (i in 1:27){
    gseList[i] <- getGEO(as.character(Series[i,1]), AnnotGPL=TRUE)
    esetFromGEO <- gseList[[i]]
    dim(exprs(esetFromGEO))
    pData(esetFromGEO)
  }
  length(gseList)
  names(gseList) <-Series[1:27,1]
  save(gseList, file="GEOSeries.Rda")
}
```

An alternative approach allowing to collect more and better iinformation is to rely on GDS information

```{r}
gdsList <- list()
# infoList <- list()
infoDatasets<- data.frame (Dataset=character(),
                           Platform=character(),
                           NumSamples=integer(),
                           colsPData = integer(),
                           col1Name = character(),
                           col1Distinct=integer(),
                           col2Name = character(),
                           col2Distinct=integer())
for (i in 1:27){
  GDSName <- as.character(Datasets[i,1])
  myGDS <- getGEO(GDSName, AnnotGPL=TRUE)
  gdsList[i]  <- myEset<- GDS2eSet(myGDS)
  phenoDat <- pData(myEset) # pData(gdsList[[i]]) 
  Platform <- Meta(myGDS)$platform
  cat("\n", i, GDSName, dim(phenoDat),  "\n" )
  cat("====================\n")
  print(head(phenoDat))
  info <- c(GDSName, Platform, 
            nrow(phenoDat), ncol(phenoDat),
            colnames(phenoDat)[2], length(unique(phenoDat[,2])),
            colnames(phenoDat)[3], length(unique(phenoDat[,3])))
  infoDatasets[i,] <- info
  # infoDF <- data.frame(matrix(rep(NA, 2*length(Meta(myGDS))), ncol=2))
  # for (i in 1:length(Meta(myGDS))){
  #   infoDF[i,1] =names(Meta(myGDS))[i]
  #   infoDF[i,2] =Meta(myGDS)[i]
  # }
  # colnames(infoDF) = c("Campo", "Descripción")
  # infoList[i]<- infoDF
}
length(gdsList)
names(gdsList) <-Datasets[1:27,1]
# names(infoList) <-Datasets[1:27,1]
save(gdsList, infoDatasets, file="GEODatasets.Rda")
```

We can use the information in `infoList` to enhance the information about the datasets.

```{r eval=FALSE}
source("https://raw.githubusercontent.com/alexsanchezpla/scripts/master/usefulFunctions/loadFromFileByName.R")
if (!exists("infoDatasets")) 
  infoDatasets<- loadFromFileByName (fileName="GEODatasets.Rda", aVarName="infoDatasets")
```

```{r}
library(dplyr)
GEODat_Enhanced <- GEOdatasets %>% 
  inner_join(infoDatasets, 
             by= c("GEO Dataset"= "Dataset"))
openxlsx::write.xlsx(GEODat_Enhanced, "GEODataset_Enhanced.xlsx")
```
